Exact asymptotic expressions for local clustering in hyperbolic random graphs \\ \TM{ I prefer not to use the term ``hyperbolic random graph". Better : KPKVB random graph or, in the title,``a hyperbolic model of complex networks"}

### 2019-01-11 introduction.tex
initial response to some comments in intro

Hyperbolic random graphs were suggested as a suitable model for complex networks by Krioukov et al.~\cite{krioukov2010hyperbolic}, exhibiting the three main characteristics commonly found in complex networks: 
\begin{enumerate}[\upshape 1)]
\item Broad degree distribution (power-law/scale-free),
\item Strong clustering (community structure),
\item Small path lengths (small world phenomenon).
\end{enumerate}
Indeed, in their seminal paper they showed that these graphs have a power-law degree distribution and exhibit strong clustering, while Bringmann et al.~\cite{bringmann2016average} showed that hyperbolic random graphs have doubly logarithmic shortest path lengths. \TM{ \RD{Remove this reference, and replace with Abdullah-Bode-Fountoulakis!!!!}
In general the proofs that GIRG = HRG are shaky at best, so I do not want to cite anything that relies on that.
Of course, in this case Nikoloas & his crew were  first anyway.
Also, regarding "distances", we probably want to mention graph diameter, studied by Kiwi+Mitsche, Friedrich+Krohmer, M\"uller + Staps.
}
Many other results regarding hyperbolic random graphs have since been established: for instance the existence of efficient routing algorithms~\cite{bringmann2017greedy}, the size of the largest component~\cite{bode2015largest},~\cite{fountoulakis2018law} and the number of cliques and the largest clique~\cite{blasius2018cliques}. \PvdH{@All: Does anyone have other suggestions for results to add?}
\TM{ Yes, of course! Everything Nikolaos & his crew, and I and my crew, and Kiwi-Mitsche have done on this model. This includes 2nd largest componnent, connectivity, spectral gap, diameter, Hamilton cycles & perfect matchings (with Markus, Nikolaos and Mitsche, in preparation but perhaps submitted around the same time as this one.).
Also there is a recent paper on arxiv by Owada+Yogeshwaran on subgraph counts.}

@@

Let $\alpha > \frac{1}{2}$, $\nu > 0$ and define $R_n = 2\log(n/\nu)$. Then the hyperbolic random graph $G_{\H,n}(\alpha, \nu)$ is defined as follows:
\begin{enumerate}
\item The vertex set if given by $n$ points $u_1, \dots, u_n$ denoted in polar coordinates $(r_i, \theta_i)$, where the angular coordinate $\theta$ is chosen uniformly from $(-\pi,\pi]$ while the radial coordinate $r$ is sampled independently according to the cumulative distribution function
\begin{equation}\label{eq:def_hyperbolic_point_distribution}
	F_{n,\alpha,\nu}(r) = \begin{cases}
		0 &\mbox{if } r < 0\\
		\frac{\cosh(\alpha r)-1}{\cosh(\alpha R_n) - 1} &\mbox{if } 0 \le r \le R_n\\
		1&\mbox{if } r > R_n
	\end{cases}
\end{equation}
\item Any two vertices $u_i=(r_i,\theta_i)$ and $u_j=(r_j,\theta_j)$ are adjacent if and only if $d_\H(u_i,u_j) \le R_n$, where $d_\H$ denotes the distance in the hyperbolic plane, i.e.
\[
	\cosh(r_i) \cosh(r_j) - \sinh(r_i) \sinh( r_j) \cos(|\theta_i-\theta_j|_{2\pi}) \le \cosh(R_n),
\]
as follows from the hyperbolic law of cosines.% rule[CITE?].
\TM{ This is only when $r_i+r_j \geq R$! Otherwise the equation does not have a solution as there is no triangle with side lengths $r_i, r_j, R$.}
\end{enumerate}

@@

\begin{theorem}[Limit for local clustering coefficient in $G_{\H,n}(\alpha,\nu)$] \label{thm:clustering_coefficient_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, and $\xi_{\alpha,\nu} = \frac{4\alpha \nu}{\pi(2\alpha -1)}$. Then,
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n} - c_\infty\right|} = 0,
\]
\TM{ As far as I can tell this is just equivalent to 
$c(G) \plim c_\infty$. I would prefer to use this more standard notation and terminology, certainly at least here, where we present our main results. }
where $c_\infty$ is defined for $\alpha \ne 1$ as
\begin{align*}
	c_\infty 
	&=\frac{2 + 4 \alpha + 13 \alpha^2 - 34 \alpha^3 - 12\alpha^4 + 24 \alpha^5}
		{16(\alpha-1)^2 \alpha (\alpha+1) (2\alpha+1)} 
		+  \frac{2^{-1 - 4 \alpha}}{(\alpha - 1)^2} \\
	&\hspace{10pt}+ \frac{(\alpha - 1/2) (B(2 \alpha, 2 \alpha + 1) + B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha))}
		{2 (\alpha - 1) (3 \alpha - 1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{2\alpha} \Gamma^\ast( - 2 \alpha, \xi_{\alpha,\nu})}{4(\alpha-1)}
		+ \frac{\xi_{\alpha,\nu}^{2\alpha + 2}\alpha (\alpha - 1/2)^2 \Gamma^\ast(- 2 \alpha - 2, \xi_{\alpha,\nu})}
		{2(\alpha-1)^2} \\
	&\hspace{10pt}- \frac{\xi_{\alpha,\nu}^{2\alpha + 1}\alpha (2\alpha - 1) \Gamma^\ast( - 2 \alpha - 1,\xi_{\alpha,\nu})}{(\alpha-1)}
		- \frac{\xi_{\alpha,\nu}^{6\alpha-2}2^{-4\alpha}(3\alpha - 1)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)^2}\\
	&\hspace{10pt}-\frac{\xi_{\alpha,\nu}^{6\alpha - 2}(\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)} \\
	&\hspace{10pt}- \frac{e^{-\xi_{\alpha,\nu}} \Gamma(2\alpha+1) 
		\left(U(2\alpha+1,1-2\alpha,\xi_{\alpha,\nu}) + U(2\alpha+1,2-2\alpha,\xi_{\alpha,\nu})\right)}{4(\alpha-1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{6\alpha - 2} \Gamma(2\alpha+1)\left( 	
		\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+2,0}{\xi_{\alpha,\nu}}
		+ \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+3,0}{\xi_{\alpha,\nu}}\right)}{4(\alpha-1)},
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty &= \frac{575 - 12 \pi^2}{576} + \frac{\eta^4(7 + \pi^2)\Gamma^\ast(-4, \eta)}{4}\\
	&\hspace{10pt}- \frac{1}{2} \int_0^1 (1 - 4z + 3z^3)\log(1-z)(z + \eta)e^{-\eta/z} \dd z\\
	&\hspace{10pt}- \int_0^1 \Li_2(z)(z^3 + \eta z^2) e^{-\eta/z} \dd z,		
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function\footnote{Note that the integrals in the expression for $c_\infty$ for $\alpha = 1$ exists: for the first one note that $1-4z+3z^2=(1-z)(1-3z)$, so the integrand can be bounded by $C(1-z)\log(1-z)$ on $[0,1)$ for some constant $C$, which can be continued continuously to the compact interval $[0,1]$ by noting that the limit for $z \rightarrow 1$ is zero, so the integrand is bounded on a bounded domain and hence, this integral is finite; for the second integral note that $\Li_2(z)$ is bounded by $\Li_2(1)$ on $[0,1]$, which is a series with well-known finite limit, so again the integrand is bounded on a bounded domain and hence the second integral is also finite.}.
\end{theorem}

@@

\TM{ I would be tempted to split this thm into two.
The first states that for $k$ fixed
$c(k) \plim c_\infty(k)$. Then we say the proof extends to ... the second thm which states that if
$1 \ll k_n \ll n^{1/(2\alpha+1)}$ then
$c(k_n) = (1+o(1)) c_\infty(k_n)$ a.a.s. \\
In fact, maybe it is good to present a separate -- and short -- proof of the constant case. The equivalence of the models for constant $k$ can be worked out in a few pages while for the sequence we need 50+ pages of super technical computations. What do you think? }

>>>>

Hyperbolic random graphs were suggested as a suitable model for complex networks by Krioukov et al.~\cite{krioukov2010hyperbolic}, exhibiting the three main characteristics commonly found in complex networks: 
\begin{enumerate}[\upshape 1)]
\item Broad degree distribution (power-law/scale-free),
\item Strong clustering (community structure),
\item Small path lengths (small world phenomenon).
\end{enumerate}
Indeed, in their seminal paper they showed that these graphs have a power-law degree distribution and exhibit strong clustering, while Bringmann et al.~\cite{bringmann2016average} showed that hyperbolic random graphs have doubly logarithmic shortest path lengths. \TM{ \RD{Remove this reference, and replace with Abdullah-Bode-Fountoulakis!!!!}
In general the proofs that GIRG = HRG are shaky at best, so I do not want to cite anything that relies on that.
Of course, in this case Nikoloas \& his crew were  first anyway.
Also, regarding "distances", we probably want to mention graph diameter, studied by Kiwi+Mitsche, Friedrich+Krohmer, M\"uller + Staps.
}
Many other results regarding hyperbolic random graphs have since been established: for instance the existence of efficient routing algorithms~\cite{bringmann2017greedy}, the size of the largest component~\cite{bode2015largest},~\cite{fountoulakis2018law} and the number of cliques and the largest clique~\cite{blasius2018cliques}. \PvdH{@All: Does anyone have other suggestions for results to add?}
\TM{ Yes, of course! Everything Nikolaos\& his crew, and I and my crew, and Kiwi-Mitsche have done on this model. This includes 2nd largest componnent, connectivity, spectral gap, diameter, Hamilton cycles \& perfect matchings (with Markus, Nikolaos and Mitsche, in preparation but perhaps submitted around the same time as this one.).
Also there is a recent paper on arxiv by Owada+Yogeshwaran on subgraph counts.}

@@

Let $\alpha > \frac{1}{2}$, $\nu > 0$ and define $R_n = 2\log(n/\nu)$. Then the hyperbolic random graph $G_{\H,n}(\alpha, \nu)$ is defined as follows:
\begin{enumerate}
\item The vertex set if given by $n$ points $u_1, \dots, u_n$ denoted in polar coordinates $(r_i, \theta_i)$, where the angular coordinate $\theta$ is chosen uniformly from $(-\pi,\pi]$ while the radial coordinate $r$ is sampled independently according to the cumulative distribution function
\begin{equation}\label{eq:def_hyperbolic_point_distribution}
	F_{n,\alpha,\nu}(r) = \begin{cases}
		0 &\mbox{if } r < 0\\
		\frac{\cosh(\alpha r)-1}{\cosh(\alpha R_n) - 1} &\mbox{if } 0 \le r \le R_n\\
		1&\mbox{if } r > R_n
	\end{cases}
\end{equation}
\item Any two vertices $u_i=(r_i,\theta_i)$ and $u_j=(r_j,\theta_j)$ are adjacent if and only if $d_\H(u_i,u_j) \le R_n$, where $d_\H$ denotes the distance in the hyperbolic plane, i.e.
\[
	\cosh(r_i) \cosh(r_j) - \sinh(r_i) \sinh( r_j) \cos(|\theta_i-\theta_j|_{2\pi}) \le \cosh(R_n),
\]
as follows from the hyperbolic law of cosines.% rule[CITE?].
\TM{ This is only when $r_i+r_j \geq R$! Otherwise the equation does not have a solution as there is no triangle with side lengths $r_i, r_j, R$.}\PvdH{I do not understand this. The hyperbolic law of cosines defines the hyperbolic cosine of the distance between points in terms of hyperbolic cosines and sines of their radii and the cosine of their relative angle:
\[
	\cosh(d_\H(u_i, u_j)) = \cosh(r_i) \cosh(r_j) - \sinh(r_i) \sinh( r_j) \cos(|\theta_i-\theta_j|_{2\pi}).
\] 
Since $\cosh$ is strictly monotonically increasing we have that $\cosh(x) \le \cosh(y) \iff x \le y$ and hence 
\begin{align*}
	&d_\H(u_i, u_j) \le R_n \\
	&\iff \cosh(d_\H(u_i, u_j)) \le \cosh(R_n) \\
	&\iff \cosh(r_i) \cosh(r_j) - \sinh(r_i) \sinh( r_j) \cos(|\theta_i-\theta_j|_{2\pi}) \le \cosh(R_n)
\end{align*}
The approximation for the hyperbolic balls you and Nicolaos derived in Lemma 28 holds only when $r_i + r_j \ge R$.}
\end{enumerate}

@@

\begin{theorem}[Limit for local clustering coefficient in $G_{\H,n}(\alpha,\nu)$] \label{thm:clustering_coefficient_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, and $\xi_{\alpha,\nu} = \frac{4\alpha \nu}{\pi(2\alpha -1)}$. Then,
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n} - c_\infty\right|} = 0,
\]
\TM{ As far as I can tell this is just equivalent to 
$c(G) \plim c_\infty$. I would prefer to use this more standard notation and terminology, certainly at least here, where we present our main results. } \PvdH{Convergence in expectation implies convergence in probability but is a much stronger notion. In addition, for integer valued sequences $a_n$ and $b_n$ we have that $\Exp{|a_n - b_n|} \to 0$ implies that $a_n = b_n$ a.a.s.. However, when they are real valued this is no longer true in general.}
where $c_\infty$ is defined for $\alpha \ne 1$ as
\begin{align*}
	c_\infty 
	&=\frac{2 + 4 \alpha + 13 \alpha^2 - 34 \alpha^3 - 12\alpha^4 + 24 \alpha^5}
		{16(\alpha-1)^2 \alpha (\alpha+1) (2\alpha+1)} 
		+  \frac{2^{-1 - 4 \alpha}}{(\alpha - 1)^2} \\
	&\hspace{10pt}+ \frac{(\alpha - 1/2) (B(2 \alpha, 2 \alpha + 1) + B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha))}
		{2 (\alpha - 1) (3 \alpha - 1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{2\alpha} \Gamma^\ast( - 2 \alpha, \xi_{\alpha,\nu})}{4(\alpha-1)}
		+ \frac{\xi_{\alpha,\nu}^{2\alpha + 2}\alpha (\alpha - 1/2)^2 \Gamma^\ast(- 2 \alpha - 2, \xi_{\alpha,\nu})}
		{2(\alpha-1)^2} \\
	&\hspace{10pt}- \frac{\xi_{\alpha,\nu}^{2\alpha + 1}\alpha (2\alpha - 1) \Gamma^\ast( - 2 \alpha - 1,\xi_{\alpha,\nu})}{(\alpha-1)}
		- \frac{\xi_{\alpha,\nu}^{6\alpha-2}2^{-4\alpha}(3\alpha - 1)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)^2}\\
	&\hspace{10pt}-\frac{\xi_{\alpha,\nu}^{6\alpha - 2}(\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)} \\
	&\hspace{10pt}- \frac{e^{-\xi_{\alpha,\nu}} \Gamma(2\alpha+1) 
		\left(U(2\alpha+1,1-2\alpha,\xi_{\alpha,\nu}) + U(2\alpha+1,2-2\alpha,\xi_{\alpha,\nu})\right)}{4(\alpha-1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{6\alpha - 2} \Gamma(2\alpha+1)\left( 	
		\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+2,0}{\xi_{\alpha,\nu}}
		+ \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+3,0}{\xi_{\alpha,\nu}}\right)}{4(\alpha-1)},
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty &= \frac{575 - 12 \pi^2}{576} + \frac{\eta^4(7 + \pi^2)\Gamma^\ast(-4, \eta)}{4}\\
	&\hspace{10pt}- \frac{1}{2} \int_0^1 (1 - 4z + 3z^3)\log(1-z)(z + \eta)e^{-\eta/z} \dd z\\
	&\hspace{10pt}- \int_0^1 \Li_2(z)(z^3 + \eta z^2) e^{-\eta/z} \dd z,		
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function\footnote{Note that the integrals in the expression for $c_\infty$ for $\alpha = 1$ exists: for the first one note that $1-4z+3z^2=(1-z)(1-3z)$, so the integrand can be bounded by $C(1-z)\log(1-z)$ on $[0,1)$ for some constant $C$, which can be continued continuously to the compact interval $[0,1]$ by noting that the limit for $z \rightarrow 1$ is zero, so the integrand is bounded on a bounded domain and hence, this integral is finite; for the second integral note that $\Li_2(z)$ is bounded by $\Li_2(1)$ on $[0,1]$, which is a series with well-known finite limit, so again the integrand is bounded on a bounded domain and hence the second integral is also finite.}.
\end{theorem}

@@

\TM{ I would be tempted to split this thm into two.
The first states that for $k$ fixed
$c(k) \plim c_\infty(k)$. Then we say the proof extends to ... the second thm which states that if
$1 \ll k_n \ll n^{1/(2\alpha+1)}$ then
$c(k_n) = (1+o(1)) c_\infty(k_n)$ a.a.s. \\
In fact, maybe it is good to present a separate -- and short -- proof of the constant case. The equivalence of the models for constant $k$ can be worked out in a few pages while for the sequence we need 50+ pages of super technical computations. What do you think? } \PvdH{I agree with the splitting suggestion. However, as I mentioned before, since $c(k_n)/c_\infty(k_n)$ is real-valued our result does not necessarily imply that $c(k_n)/c_\infty(k_n) = 1$ a.a.s.. I guess this is what you mean with $c(k_n) = (1+o(1)) c_\infty(k_n)$ a.a.s. right?}
